---
title: "Interacting with Polygon Financial Data API"
author: "David Shaw"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(httr)
library(jsonlite)
library(chron)
library(ggplot2)
library(dplyr)
API_TOKEN='qpnTdSpewzcPBpPm_jUHtmhe_peuQsi6'
```

# Requirements
The following packages must be installed using `install.packages('packageName')` and imported into environment using `library(packageName)`

 * [tidyverse](https://www.tidyverse.org/)
 * [httr](https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html)
 * [jsonlite](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html)
 * [chron](https://cran.r-project.org/web/packages/chron/chron.pdf)

Accessing the API

 * Retrieve API token at [https://polygon.io/docs/getting-started](https://polygon.io/docs/getting-started)


# API Interactions
First, place your api token in a global variable called API_TOKEN
```{r api_token, eval = FALSE}
API_TOKEN='xxxx'
```

## `tickerPrices(ticker, timespan, from, to)`
Get data for specified ticker over desired timespan between dates specified by from and to.

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
    * Forex has a ticker that begins with 'C:' followed by identifier (ex. [EUR/USD](https://finance.yahoo.com/quote/EURUSD%3DX?p=EURUSD%3DX) is C:EURUSD)
    * Crypto has a ticker that begins with 'X:' followed by an identifier (ex. [Bitcoin](https://finance.yahoo.com/quote/BTC-USD/) is X:BTCUSD)
  * `timespan`: time between measurements (minute, hour, day, week, month, quarter, year)
  * `from`: date to collect market prices from in format YYYY-mm-dd
  * `to`: date to collect market prices to in format YYYY-mm-dd
  
Return:

  * Dataframe containing volume, open/close prices, high/low prices, and number of transactions on each timespan of trading for this ticker symbol across specified dates
    
    ``` {r api1}
    tickerPrices <- function(ticker='AAPL', 
                            timespan='day', 
                            from='2021-01-01', 
                            to='2021-01-14') {
      # build url
      url <- paste('https://api.polygon.io/v2/aggs/ticker/', ticker, 
                      '/range/1/', timespan, 
                      '/', from, 
                      '/', to,
                      '/?adjusted=true&sort=asc&apiKey=', API_TOKEN, 
                      sep = '')
      
      # make api call and convert to dataframe
      raw <- GET(url)
      parsed <- fromJSON(rawToChar(raw$content))
      parsed$results
    }
    ```
    
## `marketPricesAll(market, date)`
Entire stock/forex/crypto market values for specified day.

Parameters:

  * `market`: One of `stocks`, `fx` (forex), or `crypto`
  * `date`: date to get stock market values in format YYYY-mm-dd

Return:

  * Dataframe of market prices of all symbols on specified day
  * If market is not open for that day (i.g. stock market is not open on weekends), return will be 0
  
  ```{r api2}
  marketPricesAll <- function(market = 'stocks', date='2021-01-01') {
    # build url
    marketType <- 'global'
    if(market == 'stocks') {
      marketType <- 'us'
    }
    url <- paste('https://api.polygon.io/v2/aggs/grouped/locale/', marketType,
                    '/market/', market, 
                    '/', date,
                    '?adjusted=true&apiKey=', API_TOKEN, 
                    sep = '')
    
    # make api call and convert to dataframe
    raw <- GET(url)
    parsed <- fromJSON(rawToChar(raw$content))
    parsed$results
  }
  ```

## `tickerInfo(ticker)`
Get a dataframe with information about tickers in a market.

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
    
Return:

  * Dataframe with date, and amount of dividend payment
  * If a stock does not pay dividends, an empty list will be returned
  
```{r api3}
stockDividends <- function(ticker = 'AAPL') {
  # build url
  url <- paste('https://api.polygon.io/v2/reference/dividends/', ticker,
                  '?apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```

## `marketInfo(market)`
Get a dataframe with information about tickers in a market.

Parameters:

  * `market`: One of `stocks`, `fx` (forex), or `crypto`
  
Return:

  * Dataframe containing info about active tickers in specified market
  
```{r api4}
marketInfo <- function(market = 'stocks') {
  # build url
  url <- paste('https://api.polygon.io/v3/reference/tickers?market=', market,
                  '&active=true&sort=ticker&order=asc&apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```

## `stockSplit(ticker)`
Get a dataframe with information about tickers splits in a market.

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
    
Return:

  * Dataframe with date, and ratio of split
  * If a stock has no history of splits, an empty list will be returned
  
```{r api5}
stockSplit <- function(ticker = 'AAPL') {
  # build url
  url <- paste('https://api.polygon.io/v2/reference/splits/', ticker,
                  '?apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```

## `stockFinancials(ticker, limit)`
Get a dataframe with historical financial data for a stock ticker

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
  * `limit`: number of financial reports to see
    
Return:

  * Dataframe with date, and ratio of split
  * If a stock has no history of splits, an empty list will be returned
  
```{r api6}
stockFinancials <- function(ticker = 'AAPL', limit = 5) {
  # build url
  url <- paste('https://api.polygon.io/v2/reference/financials/', ticker,
                  '?limit=', limit,
                  '&apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```


# Exploratory Analysis

## Palantir (PLTR) Analysis
Let us start by pulling year-to-date data on the ticker PLTR.
Market holidays retrieved from [here](https://www.nyse.com/markets/hours-calendars).
```{r eda_pltr1}
ticker <- 'PLTR'
timespan <- 'day'
from <- '2021-01-01'
to <- '2021-10-01'
pltr <- tickerPrices(ticker=ticker, timespan=timespan, from=from, to=to)
head(pltr)

# add a column representing dates to this dataframe
openDays <- function(from, to) {
  marketHolidays2021 <- dates(c('2021-01-01', '2021-01-18', '2021-02-15', '2021-04-02', 
                              '2021-05-31', '2021-07-05', '2021-09-06', '2021-11-25', '2021-12-24'), 
                                format="Y-M-D")
  marketOpenDates <- seq(as.Date(from), as.Date(to), by="days")
  marketOpenDates <- marketOpenDates[!is.weekend(marketOpenDates)]
  marketOpenDates <- marketOpenDates[!is.holiday(marketOpenDates, marketHolidays2021)]
}
pltr$date <- openDays(from, to)

# plot date vs closing price
pltr1 <- ggplot(pltr, aes(x=date,)) +
        geom_line(aes(y=c)) +
        ggtitle('Closing price of PLTR for Year-to-Date') + 
        theme(plot.title = element_text(hjust = 0.5, size=18, face="bold.italic")) +
        xlab('Date') + 
        ylab('Price at Close ($)')
pltr1
```

We see that PLTR achieved its high in February and has hovered around $25 since then.

Now let's look at how changes in volatility affect closing price. We first must create categorical variables to indicate if volatility has increased/decreased since the previous trading day. Likewise we will create a categorical variable to indicate if closing price has increased/decreased since the previous trading day. 
```{r eda_pltr2}
# retreive data
pltr <- tickerPrices(ticker=ticker, timespan=timespan, from=from, to=to)

# feature engineering
data <- pltr %>% 
          select(c, v) %>% 
            mutate(vIncrease = ifelse(v > lag(v), TRUE, FALSE)) %>%
              mutate(cIncrease = ifelse((c / lag(c)) > 1.05, '5% or more', 
                                  ifelse((c / lag(c)) > 1.02, '2% to 5%',
                                  ifelse((c / lag(c)) > 0.98, '-2% to 2%', 
                                  ifelse((c / lag(c)) > 0.95, '-5% to -2%',
                                         '-5% or less'))))) %>%
                mutate(order = ifelse((c / lag(c)) > 1.05, 5, 
                                    ifelse((c / lag(c)) > 1.02, 4,
                                    ifelse((c / lag(c)) > 0.98, 3, 
                                    ifelse((c / lag(c)) > 0.95, 2, 1)))))

# generate contingency table
table(data$cIncrease, data$vIncrease)
```

Above we see a breakdown of change in closing price based on volitility increases (TRUE values) and when volitility decreases (FALSE values). There seems to be correlation between volitility increasing and more extreme price changes. Let us investigate further by printing out mean and standard deviations of each category.

```{r eda_pltr3}
# feature engineering
df <- data %>% group_by(order) %>% summarise(avg=mean(v), med=median(v), sd=sd(v))
df$changeInPrice <- c('-5% or less', '-5% to -2%', '-2% to 2%', '2% to 5%', '5% or more', 'NA')

# produce dataframe to view results in organized way
df <- df %>% select(changeInPrice, avg, med, sd)
df
```

Above we see that for changes in price that exceed 5%, the average change in volitility is nearly twice that of price changes less than 5%. Additionally, the standard deviation of the mean calculation for changes above 5% is much greater than mean calculations of changes below 5% -- meaning this phenomena could be due to lack of data points for above 5% price increases.


## FAANG Analysis
FAANG stocks are known as Facebook (FB), Amazon (AMZN), Apple (AAPL), Netflix (NFLX), and Google (GOOGL). Let us analyze correlation between these stocks.

First, we need to retrieve data from API:
```{r eda_faang1}
# retrieve data
tickers <- c('AAPL', 'AMZN', 'GOOGL', 'FB', 'NFLX')
tickerPricesWrapper <- function(ticker = '') {
  timespan <- 'day'
  from <- '2021-01-01'
  to <- '2021-10-01'
  prices <- tickerPrices(ticker = ticker, timespan = timespan, from = from, to = to)
  prices$tick <- ticker

  return(prices)
}

# print out sample data
prices <- lapply(tickers, tickerPricesWrapper) %>% bind_rows()
head(prices, 10)
```

Next we will explore data using a series of plots:

* Bar plot to number/level of change for each ticker

    ```{r eda_faang2}
    # feature engineering
    faang <- prices %>% 
          select(tick, c) %>% 
            group_by(tick) %>%
              mutate(cIncrease = ifelse((c / lag(c)) > 1.05, '5% or more', 
                                  ifelse((c / lag(c)) > 1.02, '2% to 5%',
                                  ifelse((c / lag(c)) > 0.98, '-2% to 2%', 
                                  ifelse((c / lag(c)) > 0.95, '-5% to -2%',
                                         '-5% or less')))))

    # plot days increasing bar plot
    faangBar <- ggplot(faang, aes(x=tick)) +
        geom_bar(aes(fill = as.factor(cIncrease)), position = 'dodge') +
        theme(plot.title = element_text(hjust = 0.5, size=18, face="bold.italic")) +
        xlab('Date') + 
        ylab('Number of Days') + 
        labs(title = 'Level of Change vs Number of Days \n FAANG Stocks between 2021-01-01 and 2021-10-01', 
             fill = 'Level of Change')
    faangBar
    ```
  
We see that this group follows a similar distribution for level of change.

* Let's check out how volitility correlates with closing price using a scatter plot:

```{r eda_faang3}
# feature engineering
faang2 <- prices %>% 
          select(tick, c, v) %>% 
            group_by(tick) %>%
              mutate(cChange = c/lag(c)-1, vChange = v/lag(v)-1)

# plot days increasing bar plot
faangPoint <- ggplot(faang2, aes(x=vChange, y=cChange, color=tick)) +
    geom_point() +
    theme(plot.title = element_text(hjust = 0.5, size=18, face="bold.italic")) +
    xlab('Change in Volitility (Factor)') + 
    ylab('Change in Closing Price (Factor)') + 
    labs(title = 'Volitility vs Closing Price \n FAANG Stocks between 2021-01-01 and 2021-10-01', 
         color = 'Stock') + 
    xlim(-1, 2) + 
    ylim(-.1, .1)
faangPoint
```

No correlation is uncovered between changes in volitility and changes in closing price from the previous day.


## Markets Analysis

Let us explore differences between volitility in stock, forex, and crypto markets on October 4th, 2021.

```{r eda_market1}
# wrapper function
marketPricesAllVolWrapper <- function(market = 'stocks') {
  data <- marketPricesAll(market = market, date = '2021-10-04')
  data$m <- market
  return (data)
}

# retrieve data
markets <- c('stocks', 'fx', 'crypto')
marketVols <- lapply(markets, marketPricesAllVolWrapper) %>% bind_rows()

# generate box plot
marketBox <- ggplot(data = marketVols, mapping = aes(x = m, y = v, fill = m)) +
              geom_boxplot(show.legend = FALSE) +
              coord_cartesian(ylim = c(0, 10000000)) +
              theme(plot.title = element_text(hjust = 0.5, size=18, face="bold.italic")) +
                    xlab('Market Type') + 
                    ylab('Volitility') + 
                    labs(title = 'Market vs Volitility \n For the October 4th, 2021 Session')
marketBox
```


It appears that crypto has the largest volume of trades, followed by regular stock, and finally forex.


Now let's create histograms for the number of assets in each market and their volitility:

```{r eda_market2}
# wrapper function
marketPricesAllVolWrapper <- function(market = 'stocks') {
  data <- marketPricesAll(market = market, date = '2021-10-04')
  data$m <- market
  return (data)
}

# retrieve data
markets <- c('stocks', 'fx', 'crypto')
marketVols <- lapply(markets, marketPricesAllVolWrapper) %>% bind_rows()

# generate histogram
marketHist <- marketVols %>%
                  select(m, c, v) %>%
                    group_by(m) %>%
                      top_n(100, v) %>%
                        ggplot(aes(x=c, fill=m)) + 
                        geom_histogram(bins=10, show.legend = FALSE) +
                        facet_wrap(~ m, scales = "free_x") +
              theme(plot.title = element_text(hjust = 0.5, size=18, face="bold.italic")) +
                    xlab('Cosing Price (USD') + 
                    ylab('Count') + 
                    labs(title = 'Closing Prices for Top 100 Most Volitile Assets \n For the October 4th, 2021 Session')
marketHist
```

As seen above, each market's most volitile stocks tend to be the cheapest assets in that market.
