---
title: "Interacting with Polygon Financial Data API"
author: "David Shaw"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(httr)
library(jsonlite)
API_TOKEN='qpnTdSpewzcPBpPm_jUHtmhe_peuQsi6'
```

# Requirements
The following packages must be installed using `install.packages('packageName')` and imported into environment using `library(packageName)`

 * [tidyverse](https://www.tidyverse.org/)
 * [httr](https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html)
 * [jsonlite](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html)

Accessing the API

 * Retrieve API token at [https://polygon.io/docs/getting-started](https://polygon.io/docs/getting-started)


# API Interactions
First, place your api token in a global variable called API_TOKEN
```{r api_token, eval = FALSE}
API_TOKEN='xxxx'
```

## `tickerPrices(ticker, timespan, from, to)`
Get data for specified ticker over desired timespan between dates specified by from and to.

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
    * Forex has a ticker that begins with 'C:' followed by identifier (ex. [EUR/USD](https://finance.yahoo.com/quote/EURUSD%3DX?p=EURUSD%3DX) is C:EURUSD)
    * Crypto has a ticker that begins with 'X:' followed by an identifier (ex. [Bitcoin](https://finance.yahoo.com/quote/BTC-USD/) is X:BTCUSD)
  * `timespan`: time between measurements (minute, hour, day, week, month, quarter, year)
  * `from`: date to collect market prices from in format YYYY-mm-dd
  * `to`: date to collect market prices to in format YYYY-mm-dd
  
Return:

  * Dataframe containing volume, open/close prices, high/low prices, and number of transactions on each timespan of trading for this ticker symbol across specified dates
    
    ``` {r api1}
    tickerPrices <- function(ticker='AAPL', 
                            timespan='day', 
                            from='2021-01-01', 
                            to='2021-01-14') {
      # build url
      url <- paste('https://api.polygon.io/v2/aggs/ticker/', ticker, 
                      '/range/1/', timespan, 
                      '/', from, 
                      '/', to,
                      '/?adjusted=true&sort=asc&apiKey=', API_TOKEN, 
                      sep = '')
      
      # make api call and convert to dataframe
      raw <- GET(url)
      parsed <- fromJSON(rawToChar(raw$content))
      parsed$results
    }
    ```
    
## `marketPricesAll(market, date)`
Entire stock/forex/crypto market values for specified day.

Parameters:

  * `market`: One of `stocks`, `fx` (forex), or `crypto`
  * `date`: date to get stock market values in format YYYY-mm-dd

Return:

  * Dataframe of market prices of all symbols on specified day
  * If market is not open for that day (i.g. stock market is not open on weekends), return will be 0
  
  ```{r api2}
  marketPricesAll <- function(market = 'stocks', date='2021-01-01') {
    # build url
    marketType <- 'global'
    if(market == 'stocks') {
      marketType <- 'us'
    }
    url <- paste('https://api.polygon.io/v2/aggs/grouped/locale/', marketType,
                    '/market/', market, 
                    '/', date,
                    '?adjusted=true&apiKey=', API_TOKEN, 
                    sep = '')
    
    # make api call and convert to dataframe
    raw <- GET(url)
    parsed <- fromJSON(rawToChar(raw$content))
    parsed$results
  }
  ```

## `tickerInfo(ticker)`
Get a dataframe with information about tickers in a market.

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
    
Return:

  * Dataframe with date, and amount of dividend payment
  * If a stock does not pay dividends, an empty list will be returned
  
```{r api3}
stockDividends <- function(ticker = 'AAPL') {
  # build url
  url <- paste('https://api.polygon.io/v2/reference/dividends/', ticker,
                  '?apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```

## `marketInfo(market)`
Get a dataframe with information about tickers in a market.

Parameters:

  * `market`: One of `stocks`, `fx` (forex), or `crypto`
  
Return:

  * Dataframe containing info about active tickers in specified market
  
```{r api4}
marketInfo <- function(market = 'stocks') {
  # build url
  url <- paste('https://api.polygon.io/v3/reference/tickers?market=', market,
                  '&active=true&sort=ticker&order=asc&apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```

## `stockSplit(ticker)`
Get a dataframe with information about tickers splits in a market.

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
    
Return:

  * Dataframe with date, and ratio of split
  * If a stock has no history of splits, an empty list will be returned
  
```{r api5}
stockSplit <- function(ticker = 'AAPL') {
  # build url
  url <- paste('https://api.polygon.io/v2/reference/splits/', ticker,
                  '?apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```

## `stockFinancials(ticker, limit)`
Get a dataframe with historical financial data for a stock ticker

Parameters:

  * `ticker`: ticker symbol of stock/forex/crypto asset
    * Stocks have a ticker that is no more than 4 letters, all capitalized (ex. [Apple](https://finance.yahoo.com/quote/AAPL?p=AAPL&.tsrc=fin-srch) is AAPL)
  * `limit`: number of financial reports to see
    
Return:

  * Dataframe with date, and ratio of split
  * If a stock has no history of splits, an empty list will be returned
  
```{r api6}
stockFinancials <- function(ticker = 'AAPL', limit = 5) {
  # build url
  url <- paste('https://api.polygon.io/v2/reference/financials/', ticker,
                  '?limit=', limit,
                  '&apiKey=', API_TOKEN,
                  sep = '')
  
  # make api call and convert to dataframe
  raw <- GET(url)
  parsed <- fromJSON(rawToChar(raw$content))
  parsed$results
}
```


# Exploratory Analysis
placeholder

# Conclusion
placeholder
